---
title: "R_Exercise4_ShowModule3"
author: "CaAmTr"
date: "15 februari 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Module 3 Exercise

##Instructions

##Download and import the data

For this exercise, you are going to primarily be working with data from the US Bureau of Labor Statistics Consumer Expenditure Survey (CEX). You can download the data here.

- Download the SAS **2016 Interview** file. Extract the data files and keep the following two files (moving them into a project folder):

    + fmli161x.sas7bdat, which corresponds to 2016Q1 data about household characteristics and income. – mtbi161x.sas7bdat, which corresponds to 2016Q1 data about household expenditures.
    + Set the working directory.
    
- Import the two data files, calling them characteristics and expenditures. You may want to convert them to tibbles.

```{r}
library(tidyverse)
library(rio)
setwd("~/Documents/Handels/MSc in Finance/Kurs 4328/R/intrvw16/intrvw16")
 #fmli161x.sas7bdat, which corresponds to 2016Q1 data about household characteristics and income. – mtbi161x.sas7bdat, which corresponds to 2016Q1 data about household expenditures.

 characteristics <- import("fmli161x.sas7bdat")
 
 expenditures <- import("mtbi161x.sas7bdat")
```
 

##Iterate a function over column names

- Change all the column names to lower case by iterating the function tolower() over the column names of the two data frames.

```{r}
# Make column names lower case
colnames(characteristics) <- colnames(characteristics) %>% map(tolower)
colnames(expenditures) <- colnames(expenditures) %>% map(tolower)
```

## Keep specific columns

As you may’ve noticed, there’s a lot (over 800) variables in the dataset. Let’s reduce this only to the following variables:
-  **For characteristics:** newid, hh_cu_q,educ_ref, creditx,region, fincbtxm 
-  **For expenditures:** newid, cost, ref_mo, ref_yr

```{r}
# Select only desired variables in each data set
characteristics <- characteristics %>% select(newid,hh_cu_q,educ_ref,creditx, 
                                              region, fincbtxm)
expenditures <- expenditures %>% select(newid,cost,ref_mo,ref_yr)






```


##Rename columns

Rename the following variables: 
- hh_cu_q to hh_size
- fincbtxm to hh_income

```{r}
# Rename selected variables
characteristics <- characteristics %>% rename(hh_size=hh_cu_q, 
                                              hh_income=fincbtxm)

```


##Change the class of columns

Make all the variables except for newid into numeric for both data frames, using a loop or map function. Sample 80% of observations for both datasets

- To make the joins in the next step a little more interesting, first modify the datasets so that they are only a 80% sample of the full datasets.

```{r}
# Make columns except for "newid" numeric
characteristics[,-1] <- characteristics[,-1] %>% map(as.numeric)
expenditures[,-1] <- expenditures[,-1] %>% map(as.numeric)

```


##Aggregate expenditures by household

For both datasets, newid is a unique identifier for household. In the expenditures dataset, each expenditure is entered in seperately so that each household shows up many times. Using the appropriate tidyverse functions, sum up the expenditures by newid, replacing expenditures with this aggregated information.

```{r}
# Create 80% sample of data frames
characteristics <- sample_frac(characteristics, 0.8)
expenditures <- sample_frac(expenditures, 0.8)
```

##Practice different joins

Try using the different join functions covered in the module.

- In particular, first perform a traditional join that keeps all of the observations from expenditures and
the columns of expenditures and characteristics. Save the result of this join as cex_data.
- Also try a join that keeps the columns of expenditures and characteristics, but only the observations in
both datasets.
- Try a join that keeps only the columns of expenditures, with only the observations of expenditures that
are not matched in characteristics.

```{r}
# Practice different joins
cex_data <- left_join(expenditures, characteristics, by="newid")
cex_data_inner <- semi_join(expenditures, characteristics, by="newid")
cex_data_semi <- semi_join(expenditures, characteristics, by="newid")
cex_data_anti <- anti_join(expenditures, characteristics, by="newid")
```


##Use conditional statements to create indicators for region

Starting from cex_data, create indicators for each region value. You might find the unique() function helpful. Write your own simple linear regression function

```{r}
# Create regions indicators
cex_data <- cex_data %>% mutate(region1 = ifelse(region == 1,1,0),
                                region2 = ifelse(region == 2,1,0),
                                region3 = ifelse(region == 3,1,0),
                                region4 = ifelse(region == 4,1,0),
                                region4 = ifelse(is.na(region),1,0)
)


```


Finally, try your hand at writing a function. In particular, try to write a function that produces the coefficient in a linear regression. In matrix notation, the formula for βˆOLS is:
βˆOLS = (X′X)−1(X′y)
You will need some more matrix multiplication operators for this: • solve(A) yields the inverse of matrix A.
- t() provides the transpose of matrix A.
Also, remember to add a column of ones to include an intercept in the model. You can make a vector of ones by using the rep() inside of vector or matrix definition.
Once you’ve finished writing the function, try running it to produce the parameter estimates from a regression of expenditures on any of the other variables in cex_data.

```{r}
# Define ols function
my_ols <- function(indvars,depvar){
  
  # Keep only observations with no missing values for indvars and depvar
  X <- indvars[(!is.na(indvars)) & (!is.na(depvar))] 
  y <- depvar[(!is.na(indvars)) & (!is.na(depvar))]  
  
  # Create constant vector
  ones_vec <- matrix(rep(1), length(X)) 
  
  # Create matrix X equal to constant vec and indvars
  X <- cbind(ones_vec, X)     
  
  # Name constant column "constant"
  colnames(X)[1] <- "constant" 
  
  # Solve for coefficients
  beta <- solve(t(X)%*%X) %*% (t(X)%*%y) 
  colnames(beta) <- "Estimate"
  
  # Convert to data frame
  beta <- as.data.frame(beta)  
  
}

# Analysis ---------------------------------------------------------------------

# Estimate Coefficients
coeffs_schooling <- my_ols(cex_data$educ_ref, cex_data$cost)
# Display results
coeffs_schooling
```

